---
title: "Pokémon Attack Analysis"
author: "David Nguyen"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 8,
  fig.height = 5
)
# Install & load required packages
pkgs <- c("ggplot2","dplyr")
for(pkg in pkgs) if(!requireNamespace(pkg, quietly=TRUE)) install.packages(pkg)
lapply(pkgs, library, character.only=TRUE)

# Define color palette
cbPalette <- c("#E69F00","#009E73","#F0E442","#0072B2","#D55E00","#CC79A7")
````

# 1. Introduction

This report explores the **Attack** statistic across **1,008 Pokémon** from the `pokemon.csv` dataset. We will:

1. Visualize the population distribution of Attack values.
2. Compute population mean and standard deviation.
3. Count how many Pokémon fall within 1, 2, and 3 standard deviations.
4. Assess normality by comparing to the empirical rule (68–95–99.7).
5. Draw a random sample of 200 Attack values (with replacement), visualize and summarize it.
6. Discuss sampling accuracy and recommendations.

# 2. Data Loading

```{r load-data}
pokemon <- read.csv("pokemon.csv")
cat("Total Pokémon:", nrow(pokemon), "\n")
head(pokemon)
```

# 3. Population Distribution

```{r pop-histogram}
meanAttack <- mean(pokemon$Attack)
sdAttack   <- sd(pokemon$Attack)

p_pop <- ggplot(pokemon, aes(Attack)) +
  geom_histogram(binwidth = 10, fill = cbPalette[6], color = "black") +
  labs(
    title = "Distribution of Pokémon Attack Values",
    x     = "Attack",
    y     = "Frequency"
  ) +
  theme_minimal()
print(p_pop)

# Save
ggsave("attack-population.png", p_pop, width=8, height=5)
```

# 4. Population Summary

```{r pop-summary}
cat("Population Mean Attack:", round(meanAttack, 2), "\n")
cat("Population SD Attack:  ", round(sdAttack,   2), "\n")

# Empirical rule counts
within1 <- pokemon %>% filter(Attack > meanAttack - sdAttack, Attack < meanAttack + sdAttack) %>% nrow()
within2 <- pokemon %>% filter(Attack > meanAttack - 2*sdAttack, Attack < meanAttack + 2*sdAttack) %>% nrow()
within3 <- pokemon %>% filter(Attack > meanAttack - 3*sdAttack, Attack < meanAttack + 3*sdAttack) %>% nrow()

cat("Within 1 SD:", within1, sprintf("(%0.1f%%)", within1 / nrow(pokemon) * 100), "\n")
cat("Within 2 SD:", within2, sprintf("(%0.1f%%)", within2 / nrow(pokemon) * 100), "\n")
cat("Within 3 SD:", within3, sprintf("(%0.1f%%)", within3 / nrow(pokemon) * 100), "\n")
```

# 5. Normality Assessment

Based on the histogram and empirical proportions (≈66.3% within 1 SD vs. 68% expected), the Attack distribution is **slightly right-skewed** and deviates modestly from a perfect normal distribution.

# 6. Sampling Experiment

```{r sampling}
set.seed(123)
sampleAttack <- sample(pokemon$Attack, 200, replace = TRUE)
meanSample    <- mean(sampleAttack)
sdSample      <- sd(sampleAttack)

# Histogram of sample
p_samp <- ggplot(data.frame(Attack=sampleAttack), aes(Attack)) +
  geom_histogram(binwidth = 10, fill = cbPalette[6], color = "black") +
  labs(
    title = "Distribution of Sampled Pokémon Attack Values (n=200)",
    x     = "Attack",
    y     = "Frequency"
  ) +
  theme_minimal()
print(p_samp)
ggsave("attack-sample.png", p_samp, width=8, height=5)

# Sample summary
cat("Sample Mean Attack:", round(meanSample, 2), "\n")
cat("Sample SD Attack:  ", round(sdSample,   2), "\n")
```

# 7. Sampling Accuracy

The sample mean and SD differ from the population values, reflecting sampling variability. A **larger sample** (e.g., 600+) would reduce this error and better approximate the true distribution.

# 8. Conclusions

* The Attack stat exhibits a right skew and modest deviations from normality.
* Empirical rule proportions (66.3% within 1 SD) are close to theoretical expectations.
* A sample of 200 provides a rough estimate but shows noticeable variability—consider larger samples for greater accuracy.

# 9. Future Work

* Repeat this analysis for other stats (Defense, HP, etc.).
* Conduct formal normality tests (Shapiro–Wilk).
* Explore stratified sampling by Pokémon generation or type.

# 10. References

* Pokémon dataset: `pokemon.csv`
* R packages: `ggplot2`, `dplyr`
